---
title: "Gov 50 Final Project"
author: "Ee Jenn Lee"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
--- 

## Introduction
Does voter turnout in Malaysian elections exert an effect on the performance of opposition parties? In this study, I analyse data from the 14th General Election in 2018 to examine the extent to which there is a correlation between voter turnout and the voteshare received by the opposition. My hypothesis is that higher voter turnout is associated with more votes for the opposition, whereas lower voter turnout benefits the incumbent party. This is because individuals who do not vote are more likely to oppose the incumbent party as dissatisfaction with the administration is often expressed as voter apathy. Such individuals may see voting as validating electoral fraud and systems of gerrymandering and malapportionment, for instance. Moreover, specifically in Malaysia, there are incentives, financial or otherwise, for supporters of the incumbent regime to turn out to vote. When I run my regression of opposition voteshare on voter turnout, a positive, significant coefficient would indicate support for my hypothesis.



```{r DATA CLEANING}
library(tidyverse)

ge14 <- read_csv("~/Desktop/Gov 50/Election-Results-2018 - Parlimen_Results_By_Candidate.csv")
ge14[is.na(ge14)] = 0
ge14$`% of total Votes` <- gsub("%", "", ge14$`% of total Votes`)
ge14$`% of total Votes` <- as.numeric(ge14$`% of total Votes`)
# ge14

turnout <- read_csv("~/Desktop/Gov 50/General-Election-Data-master/Voter Turnout GE12, GE13, GE14 - Sheet1.csv")

turnout$`GE 13 VOTER TURNOUT (2013) %` <- gsub("%", "", turnout$`GE 13 VOTER TURNOUT (2013) %`)
turnout$`GE 13 VOTER TURNOUT (2013) %` <- as.numeric(turnout$`GE 13 VOTER TURNOUT (2013) %`)

turnout$`GE 14 VOTER TURNOUT (2018) %` <- gsub("%", "", turnout$`GE 14 VOTER TURNOUT (2018) %`)
turnout$`GE 14 VOTER TURNOUT (2018) %` <- as.numeric(turnout$`GE 14 VOTER TURNOUT (2018) %`)

turnout$`GE 12 VOTER TURNOUT (2008) %` <- gsub("%", "", turnout$`GE 12 VOTER TURNOUT (2008) %`)
turnout$`GE 12 VOTER TURNOUT (2008) %` <- as.numeric(turnout$`GE 12 VOTER TURNOUT (2008) %`)
# turnout
```


```{r Joining Datasets & Data Cleaning}
library(dplyr)
ge14 <- full_join(turnout, ge14, by = c("PARLIAMENT CODE" = "Seat ID"))

ge14 <- ge14 |>
  transform(
  party_switch,
  party_switch = replace(party_switch, `Candidate Name` == "AMIN AHMAD", 1
                         )) |>
  transform(
  party_switch,
  party_switch = replace(party_switch, PARLIAMENT == "PADANG BESAR", 0)
)  |>
  transform(
  party_switch,
  party_switch = replace(party_switch, PARLIAMENT == "BATU", 0)
) |>
  transform(
  party_switch,
  party_switch = replace(party_switch, PARLIAMENT == "PENAMPANG", 0))
  
# 
# ge14
```




```{r Average Voter Turnout for Incumbent vs Party Switch}
hist <- ge14 |> 
  select(PARLIAMENT, party_switch, GE.14.VOTER.TURNOUT..2018..., Candidate.Win) |>
  filter(Candidate.Win == 1) |>
  group_by(party_switch) |>
  summarize(mean_turnout = mean(GE.14.VOTER.TURNOUT..2018..., na.rm = TRUE
)) |>
  mutate(party_switch = if_else(party_switch == 0, "No", "Yes"))
                                                                       
knitr::kable(hist, digits = 4)
```

```{r Plot}
library(ggplot2)

ggplot(hist, aes(x = party_switch,
                 y = mean_turnout))+
  geom_col(fill = "dodgerblue") + 
  labs(title = "Party switch against voter turnout",
    x = "Party Switch",
      y = "Mean Turnout Percentage(%)") +
  coord_cartesian(ylim=c(75,85)) 
```
## Describing my data
My sample is comprised of the 222 parliamentary seats that were contested in the 2018 Malaysian general elections. My unit of analysis is a parliamentary seat. The explanatory variable of interest is voter turnout as a percentage of the voting-age population, which is a number out of 100. My outcome variable is the cumulative voteshare received by opposition parties (i.e. those that did not occupy that seat in the previous election), which is a number out of 100. I derived this number by taking 100 - the voteshare received by the incumbent party. Because of the nature of my data, when the incumbent party did not retain their seat, I had to manually input this information. 

This is an observational study. Neither of the three observational study research designs that were covered in lecture apply to my study as there is no treatment being applied and no treatment and control groups; the analysis is merely investigating the correlation between two variables. 

However, this plot is an example of a cross-sectional design. The plot summarizes the independent variable (voter turnout) more so than the dependent variable (opposition voteshare), but it provides useful insights into the dependent variable in a different form as it looks at whether opposition parties are able to win seats. Party_switch = no means that the same party retains its seat and party_switch = yes means that a different party won the seat. The plot shows mean voter turnout when incumbents win versus when they do not, and preliminarily supports my hypothesis as it shows that higher voter turnout is associated with opposition party wins.


```{r}
ge14_partyswitch <- ge14 |>
    select(STATE, `PARLIAMENT.CODE`, `PARLIAMENT`, party_switch, GE.14.VOTER.TURNOUT..2018...) |>
filter(party_switch == 1) 

voteshare_opposition <- c(65.74, 70.86, 71.75, 71.86, 66.92, 66.7, 66.88, 64.53, 70.81, 65.88, 55.24, 58.33, 61.81, 71.18, 71.77, 96.48, 61.65, 64.43, 89.65, 57.02, 59.63, 67.86, 64.26, 78.39, 83.7, 84.74, 82.48, 66.97, 60.9, 55.65, 69.81, 61.34, 59.85, 58.99, 51.69, 58.47, 61.03, 60.19, 63.43, 58.85, 59.01, 73.73, 64.87, 62.31, 69.48, 53.88, 58.42, 67.7, 61.78, 52.59, 58.1, 68.26, 54, 64.07, 60.62, 81.65, 61, 58.55, 56.71, 56.95, 67.18, 52.18, 55.28, 51.11)
ge14_partyswitch$voteshare_opposition <- voteshare_opposition
# ge14_partyswitch

ge14_noswitch <- ge14 |>
filter(party_switch == 0) |>
  filter(Candidate.Win == 1) |>
  select(STATE, `PARLIAMENT.CODE`, `PARLIAMENT`, party_switch, GE.14.VOTER.TURNOUT..2018..., `X..of.total.Votes`) |>
  mutate(voteshare_opposition = 100 - `X..of.total.Votes`) |>
  select(-X..of.total.Votes)
# ge14_noswitch

ge14_turnout_voteshare <- rbind(ge14_partyswitch, ge14_noswitch)
ge14_turnout_voteshare <- ge14_turnout_voteshare[-c(73, 106),]
# ge14_turnout_voteshare

```


```{r}
## Regression on opposition vote share and voter turnout
fit <- lm(voteshare_opposition~GE.14.VOTER.TURNOUT..2018..., data = ge14_turnout_voteshare)
coef(fit)

library(broom)
tidy(fit) |>
  knitr::kable(caption = "Coefficient Estimates for Opposition Voteshare on Voter Turnout",
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,3,3,3,3))

## Multiple regression controlling for state differences
fit_multi <- lm(voteshare_opposition~GE.14.VOTER.TURNOUT..2018... + STATE, data = ge14_turnout_voteshare)
tidy(fit_multi) |>
  knitr::kable(caption = "Coefficient Estimates for Opposition Voteshare on Voter Turnout (controlling for state effects)",
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"),
        digits = c(0,3,3,3,3))
```
```{r}
## Scatterplot of regression
fig <- ggplot(ge14_turnout_voteshare,
       aes(x = GE.14.VOTER.TURNOUT..2018..., 
           y = voteshare_opposition)) +
  geom_point(color = "steelblue1") +
  geom_smooth(method = "lm", se = FALSE, color = "indianred1", size = 1.5) +
  labs(x = "GE 14 Voter Turnout",
       y = "Vote share received by the opposition",
       title = "Voter turnout vs opposition performance") 
fig
ggsave("fig.png", fig, width = 10, height = 5)

```
## Results
Linear regression:
The correlation coefficient of opposition voteshare on voter turnout is 0.742, indicating that when voter turnout increases by 1%, the opposition voteshare increases by 0.742%. The p-value for this regression is 0.001, which is less than 0.05 and hence is statistically significant. 

Because it is possible that this finding is due to confounding factors, I also ran a multiple regression to account for fixed state effects:
The correlation coefficient of opposition voteshare on voter turnout, controlling for unobserved variation between states, is 1.080, indicating that when voter turnout increases by 1%, the opposition voteshare increases by 1.08%. The p-value is 0.004, which is statistically significant. There is, however, a lot of variation between states, as some states experience a very large negative correlation between voter turnout and opposition voteshare and some experience a very large positive correlation, which produces on aggregate the small positive effect that we observe. Hence, I do not believe that the coefficient represents a causal effect even though it is statistically significant.

## Conclusion
In conclusion, the results do show that there is a positive and statistically significant relationship between voter turnout and opposition performance, but I do not believe that the coefficient represents a causal effect for two reasons.

Firstly, as mentioned above, because of the significant variation between states. It appears that there is no general nation-wide effect of voter turnout on opposition performance; if there was, we would expect the results for each state to be roughly the same, i.e. a small, positive effect. The fact that the correlation coefficients for 8 states are negative, 2 of them being very negative (~ -19 and -17) does not provide much confidence in our findings. 

Secondly, because this is an observational study and not a randomized control trial, concluding that there is a causal effect would require accounting for all possible confounders, which is not really possible. Aside from fixed state effects, I could have improved the study by adding more covariates, such as: which party was the incumbent party (as some have more political power and financial resources than others), whether the seat was a swing seat or a party stronghold, voter demographics etc. However, it is impossible to account for all possible covariates, which is one significant limitation of a study like this.

```



