---
title: "Gov 50 Final Project"
author: "Ee Jenn Lee"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
--- 



## Project thoughts
I am interested in exploring data related to the upcoming Malaysian elections. Making predictions about election outcomes would be cool, but it might also be outside my skill set. I'm particularly interested in exploring the potential effect of a recent law lowering the voting age to 18 on election outcomes, and would look at data sets from youth opinion surveys and state-level elections to explore the role of youth in Malaysian politics. 

## Research proposal
Does voter turnout in Malaysian elections exert an effect on whether the incumbent party is able to remain in power? In this study, I plan to examine state-level data of past Malaysian elections to examine the extent to which there is a correlation between voter turnout and the party that wins. My hypothesis is that lower voter turnout benefits the incumbent party, because dissatisfaction with the administration is often expressed as voter apathy and specifically in Malaysia, supporters of the regime are incentivized to vote by rewards, financial or otherwise. My sample consists of state-level data from the 14th General Election in 2018 [ ]. My explanatory variable is voter turnout as a percentage of the voting-age population. My outcome variable is the party that wins the local election, which will be coded discretely and numerically. A significant positive correlation (generated by a multiple regression analysis that adjusts for potential confounders) would support my hypothesis. A negligible or negative correlation would provide evidence against my hypothesis.

```{r}
library(tidyverse)
turnout <- read_csv("~/Desktop/Gov 50/General-Election-Data-master/Voter Turnout GE12, GE13, GE14 - Sheet1.csv")
turnout
ge14 <- read_csv("~/Desktop/Gov 50/Election-Results-2018 - Parlimen_Results_By_Candidate.csv")
ge14[is.na(ge14)] = 0
ge14$`% of total Votes` <- gsub("%", "", ge14$`% of total Votes`)
ge14$`% of total Votes` <- as.numeric(ge14$`% of total Votes`)

ge14

turnout$`GE 13 VOTER TURNOUT (2013) %` <- gsub("%", "", turnout$`GE 13 VOTER TURNOUT (2013) %`)
turnout$`GE 13 VOTER TURNOUT (2013) %` <- as.numeric(turnout$`GE 13 VOTER TURNOUT (2013) %`)

turnout$`GE 14 VOTER TURNOUT (2018) %` <- gsub("%", "", turnout$`GE 14 VOTER TURNOUT (2018) %`)
turnout$`GE 14 VOTER TURNOUT (2018) %` <- as.numeric(turnout$`GE 14 VOTER TURNOUT (2018) %`)

turnout$`GE 12 VOTER TURNOUT (2008) %` <- gsub("%", "", turnout$`GE 12 VOTER TURNOUT (2008) %`)
turnout$`GE 12 VOTER TURNOUT (2008) %` <- as.numeric(turnout$`GE 12 VOTER TURNOUT (2008) %`)
turnout
```


```{r}
library(dplyr)

ge14 <- full_join(turnout, ge14, by = c("PARLIAMENT CODE" = "Seat ID"))
ge14
```


```{r}
hist <- ge14 |> 
  select(`PARLIAMENT`, party_switch, `GE 14 VOTER TURNOUT (2018) %`, `Candidate Win`) |>
  filter(`Candidate Win` == 1) |>
  group_by(party_switch) |>
  summarize(mean_turnout = mean(`GE 14 VOTER TURNOUT (2018) %`, na.rm = TRUE
)) |>
  mutate(party_switch = if_else(party_switch == 0, "No", "Yes"))

hist
```

```{r}
library(ggplot2)

ggplot(hist, aes(x =party_switch,
                 y = mean_turnout))+
  geom_col(fill = "dodgerblue") + 
  labs(title = "Party switch against voter turnout",
    x = "Party Switch",
      y = "Mean Turnout Percentage(%)") +
  coord_cartesian(ylim=c(75,85)) 
```

## For non-switch, run regression on % of votes received by opposition (the other candidates for whic candidate win = 0) vs voter turnout 
## If party switch = 1,
```{r}
ge14 <- ge14 |> 
  transform(
  party_switch,
  party_switch = replace(party_switch, `Candidate Name` == "AMIN AHMAD", 1
                         )) |>
  transform(
  party_switch,
  party_switch = replace(party_switch, PARLIAMENT == "PADANG BESAR", 0)
)  |>
  transform(
  party_switch,
  party_switch = replace(party_switch, PARLIAMENT == "BATU", 0)
)  |>
  transform(
  party_switch,
  party_switch = replace(party_switch, PARLIAMENT == "PENAMPANG", 0)
) 
ge14

```
```{r}
ge14_partyswitch <- ge14 |>
    select(STATE, `PARLIAMENT.CODE`, `PARLIAMENT`, party_switch, GE.14.VOTER.TURNOUT..2018...) |>
filter(party_switch == 1) 

voteshare_opposition <- c(65.74, 70.86, 71.75, 71.86, 66.92, 66.7, 66.88, 64.53, 70.81, 65.88, 55.24, 58.33, 61.81, 71.18, 71.77, 96.48, 61.65, 64.43, 89.65, 57.02, 59.63, 67.86, 64.26, 78.39, 83.7, 84.74, 82.48, 66.97, 60.9, 55.65, 69.81, 61.34, 59.85, 58.99, 51.69, 58.47, 61.03, 60.19, 63.43, 58.85, 59.01, 73.73, 64.87, 62.31, 69.48, 53.88, 58.42, 67.7, 61.78, 52.59, 58.1, 68.26, 54, 64.07, 60.62, 81.65, 61, 58.55, 56.71, 56.95, 67.18, 52.18, 55.28, 51.11)
ge14_partyswitch$voteshare_opposition <- voteshare_opposition
ge14_partyswitch

```
```{r}
fit_partyswitch <- lm(voteshare_opposition~GE.14.VOTER.TURNOUT..2018..., data = ge14_partyswitch)
coef(fit_partyswitch)

library(broom)
augment(fit_partyswitch) |> head()

ggplot(ge14_partyswitch,
       aes(x = GE.14.VOTER.TURNOUT..2018..., 
           y = voteshare_opposition)) +
  geom_point(color = "steelblue1") +
  geom_smooth(method = "lm", se = FALSE, color = "indianred1", size = 1.5) +
  labs(x = "GE 14 Voter Turnout",
       y = "Vote share received by the opposition",
       title = "Voter turnout vs opposition performance for flipped seats") 

fit_partyswitch_multi <- lm(voteshare_opposition~GE.14.VOTER.TURNOUT..2018... + STATE, data = ge14_partyswitch)
coef(fit_partyswitch_multi)
```
```{r}
ge14_noswitch <- ge14 |>
filter(party_switch == 0) |>
  filter(Candidate.Win == 1) |>
  select(STATE, `PARLIAMENT.CODE`, `PARLIAMENT`, party_switch, GE.14.VOTER.TURNOUT..2018..., `X..of.total.Votes`) |>
  mutate(voteshare_opposition = 100 - `X..of.total.Votes`)
ge14_noswitch


fit_noswitch <- lm(voteshare_opposition~GE.14.VOTER.TURNOUT..2018..., data = ge14_noswitch)
coef(fit_noswitch)

augment(fit_noswitch) |> head()

ggplot(ge14_noswitch,
       aes(x = GE.14.VOTER.TURNOUT..2018..., 
           y = voteshare_opposition)) +
  geom_point(color = "steelblue1") +
  geom_smooth(method = "lm", se = FALSE, color = "indianred1", size = 1.5) +
  labs(x = "GE 14 Voter Turnout",
       y = "Vote share received by the opposition",
       title = "Voter turnout vs opposition performance for seats won by incumbents")

fit_noswitch_multi <- lm(voteshare_opposition~GE.14.VOTER.TURNOUT..2018... + STATE, data = ge14_noswitch)
coef(fit_noswitch_multi)
```
## Questions for section: 
## where do I go from here? what additional analysis shohuld i do? should i combine my incumbent and party switch data sets?
## hold states fixed?
## is this a ggood way to test the relationship between my variables?
## shhould i run a hypothesis test to show hhow likely it would be that i get these results?(a brief (one paragraph) concluding section that summarizes your results, assesses the extent to which you find support for your hypothesis, describes limitations of your analysis and threats to inference, and states how your analysis could be improved (e.g., improved data that would be useful to collect))


```